Paper

### 2018-01-05 Paper.Rmd
...

@@

globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
globalCN.plot = globalCN.plot + theme(legend.position = 'none')

nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
#globalCNBEdf = subset(globalCNdf, Pathology == 'BE') %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

globalCN.plot 
globalBECN.plot

@@

a = sample(subset(patient.info, Status == 'NP' )$Samplename, 1)
b = sample(subset(patient.info, Status == 'P' )$Samplename, 1)

cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual samples")
cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both), type='bar') + labs(title="Individual samples")

@@

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
folds = 10; splits = 5 
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))
```

@@

mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))

@@

classification.rate = add_row(classification.rate, 'model'='All samples', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
```

@@

```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=F}
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
```

```{r, echo=F, fig.height=5, fig.width=5}
mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('No HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]])
                              )
```

@@

```{r, echo=F, fig.height=5, fig.width=5}
mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('No LGD/HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
```

@@

```{r, echo=F, warning=F, message=F,fig.height=5, fig.width=5}
#classification.rate$x = 1:nrow(classification.rate)

ggplot(classification.rate, aes(model, mean, group=model, color=model)) + 
  geom_text( aes(label=round(mean, 3)), hjust=-0.5, size=5, color='gray39', show.legend = F) +
  geom_text( aes(y=mean+sme*2, label=paste(all.coef, "\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
  geom_point(size=4) +
  labs(x='', y='Classificaiton rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none')

@@

```{r, echo=F, warning=F, message=F}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, performance.at.1se)

@@

@@

```{r leaveoneout, echo=F, message=F, warning=F}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = coefs

df = as.data.frame(performance.at.1se)
ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='performance at lambda.1se', x='', title=paste('Performance for nested leave one out, at alpha=',select.alpha,sep='')) +
  plot.theme

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$Prediction))

@@

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, ci=T, of='thresholds')
roc$model = 'CN Model'

grid.arrange(roc.plot(roc50,'Probability threshold = 0.5'), roc.plot(roc, 'Probability threshold calculated by ROC'), ncol=2)

roc.plot(roc, 'Copy number model')

@@

@@

```{r echo=F, fig.height=6, fig.width=8}
get.roc.stat<-function(roc) {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )

  sens = c(pROC::coords(roc, 'best')[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, 'best')[['specificity']], conf.level=0.95))
  spec = c(pROC::coords(roc, 'best')[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, 'best')[['sensitivity']], conf.level=0.95))

@@

m = rbind.data.frame(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC','HGD Only','Dysplasia (HGD & LGD)','CN Model'), ordered = T)

@@

aucs = do.call(rbind, lapply(rocList,function(r) cbind.data.frame('AUC'=r$auc*100,'model'=r$model,'Specificity'=pROC::coords(r,'best')[['specificity']], 'Sensitivity'=pROC::coords(r,'best')[['sensitivity']])))

df = do.call(rbind, lapply(rocList, function(r) 
              cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model) 
              ))

ggplot(df, aes(Specificity, Sensitivity,color=model,fill=model)) + geom_line(show.legend = F) + scale_x_reverse() + 
    geom_label_repel(data=aucs, aes(label=paste(model,'\nAUC:', sprintf("%.2f",AUC), '%', sep='')), nudge_x=.25, color='white', show.legend = F) + 
  labs(title='ROC Curves', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

dodge = position_dodge(width=0.9)
ggplot(melt(m, measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=0, vjust=-1 ), position=dodge, color='white')  +
  scale_fill_manual(values=c('purple3','dodgerblue3','firebrick3'), name='') + plot.theme + 
  theme(legend.position = 'bottom') + labs(x='', y='Percent', title='ROC Comparison') 
```

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
  )

@@

back$Pathology = as.character(back$Pathology)
back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','IMC/HGD'), ordered = T)

@@

```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
cuts = seq(0,1,0.1)
#cuts = quantile(back$Prediction, probs=seq(0,1,1/5))

@@

qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction), 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))

ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')

pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

@@

@@

#annotation.table = as.data.frame(annotation.table)
#rownames(annotation.table) = annotation.table$Risk
#annotation.table$Risk = NULL

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = c('green3','orange2','red3'), col=NA, alpha=0.2),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="black", fontface=4L)),
  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
  padding=unit(c(5,5),'mm'))

grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )

>>>>

riskCols = brewer.pal(11, "RdYlBu")[c(10, 5, 1)]
```

@@

globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
#globalCN.plot = globalCN.plot + theme(legend.position = 'none')

## No HGD
`%nin%` = Negate(`%in%`)

nhgd.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %nin% c('HGD','IMC'))$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %nin% c('HGD','IMC'))$variable)),')', sep=''))
globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %nin% c('HGD','IMC')), labeller(Status=nhgd.status, chr=label_value)) + labs(title='Excluding HGD/IMC')

## NDBE
nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

globalCN.plot 
globalnHGDCN.plot
globalBECN.plot

@@

a = sample( subset(subset(patient.info, Patient == 57),  Endoscopy.Year == min(Endoscopy.Year) )$Samplename, 1)
b = sample( subset(subset(patient.info, Patient == 76), Endoscopy.Year == min(Endoscopy.Year) & Pathology %nin% c('HGD','IMC'))$Samplename, 1)

ind.status <- c('Non-Progressor'='Patient 57 (n=1)',
              'Progressor'='Patient 76 (n=1), excl. HGD/IMC')
cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(Status = ind.status), type='bar') + labs(title="Individual samples")
```

@@

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
folds = 10; splits = 5 
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))

```{r echo=F, fig.height=5, fig.width=5}
library(ggfortify)
autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme
```

@@

mp = model.performance(performance.at.1se, coefs, lims=c(0.7,0.8))
mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))

@@

classification.rate = add_row(classification.rate, 'model'='All samples', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

@@

```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=T}
rm(performance.at.1se, coefs, mp)

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs, lims=c(0.69,0.8))
mp$plot + labs(title='No HGD/IMC samples')

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]])
                              )

rm(performance.at.1se, coefs, mp)
```

#### LOO minus HGD

```{r loonohgd, echo=F, message=F, warning=T}

file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=F)
} else {
  stop(paste("No file", file, "found"))
}
```

```{r, echo=F, message=F, fig.width=5, fig.height=5}

  pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)

  perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))

  ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')

  rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')

```{r, echo=F, fig.width=6, fig.height=6}
ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme

cuts = seq(0,1,0.1) 
pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))

noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)

noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

noHGD.pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

noHGD.pred.confidence$Risk = 'Moderate'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))

noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() + geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')

# Predict HGD samples
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)

  fitNOHGD = models[[select.alpha]]

  lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
  hgd.df = dysplasia.df[hgd.samples$Samplename,]

  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
  hgd.samples$Prediction = pred

  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )

  rm(no.hgd.plots, coefs, performance.at.1se, models)
}

rm(plots, performance.at.1se,coefs, nzcoefs, fits)
```

@@

#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs, lims=c(0.69,0.8))
mp$plot + labs(title='No LGD/HGD/IMC samples')

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp)
```

@@

```{r, echo=F, warning=F, message=F, fig.width=6, fig.height=6}
#classification.rate$x = 1:nrow(classification.rate)

ggplot(classification.rate, aes(model, mean, group=model, color=model)) + 
  geom_text( aes(label=round(mean, 3)), hjust=-0.5, size=5, color='gray39', show.legend = F) +
  geom_text( aes(y=mean+sme*2, label=paste(all.coef, "\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
  geom_point(size=4) +
  labs(x='', y='Classificaiton rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') + 
  ylim(0.7,0.8)

@@

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, performance.at.1se)

@@

```{r nohgd_coefs, message=F, warning=F, echo=F}
setdiff(ch$label, rownames(nohgd.coefs))

@@

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

df = as.data.frame(performance.at.1se)
ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=5, fig.width=5}
preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$Prediction))

@@

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'

# To find the auc at a given spec or sens
#pROC::auc(roc, partial.auc=c(0.99,1), partial.acu.focus="sp", partial.auc.correct=T)

df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))

thresholds = 
  rbind.data.frame(
#    est=t(round(pROC::coords(roc, "best"),2)),
    max=t(round(pROC::coords(roc, 0.8),2)),
    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
    'All samples'=t(round(pROC::coords(roc, 0.6),2))
)
thresholds$label=rownames(thresholds)

ggplot() + geom_line(data=df, aes(specificity, sensitivity)) + scale_x_reverse() + 
  geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
  geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
    paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
  labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

roc.plot(roc, 'Probability threshold calculated by ROC')

roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
pROC::coords(rocNOHGD, "best")
pROC::coords(rocNOHGD, 0.6)
pROC::auc(rocNOHGD, partial.auc=c(0.89,0.9), partial.auc.focus='spec', partial.auc.correct=T)

@@

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}

pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% summarise(
  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
)

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best')
#roc.plot(rocEndoM)
rocEndoM$model = 'Per Endoscopy Max'

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best')
#roc.plot(rocEndoA)
rocEndoA$model = 'Per Endoscopy Avg'

lastEndo = do.call(rbind, lapply(pg.samp, function(pt) {
 cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[nrow(pt), 'PID'])$Prediction))
}))

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
pROC::coords(rocLastEndo, 'best')
#roc.plot(rocLastEndo)
rocLastEndo$model = 'Last endoscopy'

firstEndo = do.call(rbind, lapply(pg.samp, function(pt) {
 cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[1, 'PID'])$Prediction))
}))

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
pROC::coords(rocFirstEndo, 'best')
#roc.plot(rocFirstEndo)
rocFirstEndo$model = 'First endoscopy'

### Per Patient ROC

```{r echo=F, fig.height=4, fig.width=4}

ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise( 'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtM, 'best')
#roc.plot(rocPtM)
rocPtM$model = 'Per Patient Max'

rocPtA = pROC::roc(Status ~ avg, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtA, 'best')
#roc.plot(rocPtA)
rocPtA$model = 'Per Patient Avg'

@@

```{r echo=F, fig.height=6, fig.width=8}
roc$model = "Per Sample"
multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons')
```

```{r echo=F, fig.height=8, fig.width=6}
get.roc.stat<-function(roc, x='best') {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )

  sens = c(pROC::coords(roc, x)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x)[['specificity']], conf.level=0.95))
  spec = c(pROC::coords(roc, x)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x)[['sensitivity']], conf.level=0.95))

@@

get.auc.at<-function(roc, foc='spec', p=c(1,0.99)) {
# To find the auc at a given spec or sens
  pROC::auc(roc, partial.auc=p, partial.auc.focus=foc, partial.auc.correct=T)

get.auc.at(p53roc)
get.auc.at(hgdpathroc)
get.auc.at(pathroc)
get.auc.at(roc)

m = rbind.data.frame(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc, 0.6))
m$model = factor(m$model, levels=c('p53 IHC','HGD Only','Dysplasia (HGD & LGD)','CN Model'), ordered = T)

@@

dodge = position_dodge(width=0.9)
ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=0, vjust=-1 ), position=dodge, color='white')  +
  scale_fill_manual(values=c('purple3','dodgerblue3','firebrick3'), name='') + plot.theme + 
  theme(legend.position = 'none') + labs(x='', y='Percent', title='ROC Comparison') 
```

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)

@@

back$Pathology = as.character(back$Pathology)
#back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)

@@

```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
cuts = seq(0,1,0.1)

@@

qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
#ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')

pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

@@

pred.confidence = bind_cols(pred.confidence, 
                            data.frame(ci.low=qbeta(0.025, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5),
                                      ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))

@@

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = riskCols, col=NA, alpha=0.2),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="black", fontface=4L)),
  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
  padding=unit(c(5,5),'mm'))

#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
```

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() + geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none')
```

```{r probhist, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
    geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
    geom_text(data=annotation.table %>% group_by(Risk, min,max) %>% summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') + theme(legend.position = 'none')
```

```{r indP, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
back = back %>% group_by(Risk) %>% mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )
